---
title: "The effects of naproxen sodium on the growth, reproduction, survival, and feeding of a freshwater pond snail"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
# Packages and data reading
```{r package_data}
library(tidyverse) # formatting and tidying data
library(glmmTMB) # performing glmms
library(DHARMa) # model validation
library(survival) # survival analysis
library(ggplot2) # creating figures
library(scales)
library(dplyr)
library(emmeans)
library(car)
library(survminer)

Data <- read.csv("Naproxen_Physa_rawdat.csv") # read in data
color_pal <- rev(viridis_pal()(8))[c(2,4,6,8)]
```



# SnailGrowth

Snails were measured from apex to base with calipers once a week for the month long experiment.  The data was analyzed using a linear mixed effects model with snail as a random effect and Week and Treatment as fixed effects.  

```{r Growth}
Growth_data <- Data %>%
  dplyr::select(Treatment,Snail,Week,Length,Status) %>%
  filter(Status=="alive") %>% 
  mutate(Week = paste("Week",Week,sep="")) %>%
  group_by(Treatment,Snail) %>%
  pivot_wider(names_from=Week,values_from=Length) %>%
  mutate(Growth1 = (Week1-Week0),
         Growth2 = (Week2-Week1),
         Growth3 = (Week3-Week2),
         Growth4 = (Week4-Week3)
  ) %>%
  dplyr::select(-c(Week0,Week1,Week2,Week3,Week4))%>%
  pivot_longer(cols = Growth1:Growth4,names_to="Week",values_to="Growth") %>%
  mutate(Week = str_replace(Week, 'Growth', ""),
         Week = as.numeric(Week))


Week1 <- Growth_data %>% filter(Week==1)
mean(Week1$Growth)

OtherWeeks <- Growth_data %>% filter(Week!=1)
mean(OtherWeeks$Growth,na.rm=TRUE)


# Fitting the model
mod_growth <- glmmTMB(Growth~Week*Treatment+I(Week^2) + (1|Snail), data=Growth_data) 
summary(mod_growth)
Anova(mod_growth)

# Model validation
simulationOutput <- simulateResiduals(fittedModel = mod_growth, plot = T) 
## Note: Including the quadratic term in the model greatly reduced but did not completely eliminate issues with quantile deviations

# Plotting the data
Growth_data$TreamentPlot <- factor(Growth_data$Treatment,levels=expression())

growth_plot <- ggplot(Growth_data,aes(x=Week,y=Growth,fill=Treatment,group=interaction(Treatment,Week)))+
  geom_boxplot()+
  theme_classic()+
  ylab("Growth rate (mm)")+
  scale_fill_manual(values=color_pal,
                    labels = c(expression(paste(0," ",mu,"g/L")),
                               expression(paste(100," ",mu,"g/L")),
                               expression(paste(1000," ",mu,"g/L")),
                               expression(paste(10000," ",mu,"g/L")))
                    )+
 # theme(legend.position=c(.9,.8))
  theme(legend.position="none")
```

# Reproduction (number of egg masses)

We recorded the number of eggs per snail each week.  Because these are counts, we used a Poisson glmm to analyze the data.  

## Analysis

```{r Reproduction}
EggData <- Data %>% filter(Week!=0) %>% filter(Status=="alive")


Week1 <- EggData %>% filter(Week==1)
mean(Week1$Egg.Sacs)

Week4 <- EggData %>% filter(Week==4)
mean(Week4$Egg.Sacs)

# glmm
model_eggs <- glmmTMB(Egg.Sacs ~ Week*Treatment + (1 | Snail),
                data = EggData,
                family="poisson")

summary(model_eggs)
Anova(model_eggs)

# model validation
simulationOutput <- simulateResiduals(fittedModel = model_eggs, plot = T)

```

## Making the figure
```{r Reproduction_fig}
reproduction_plot <- ggplot(EggData,aes(x=Week,y=Egg.Sacs,fill=Treatment,group=interaction(Treatment,Week)))+
  geom_boxplot()+
  ylab("Number of egg sacs")+
  theme_classic()+
    scale_fill_manual(values=color_pal,
                    labels = c(expression(paste(0," ",mu,"g/L")),
                               expression(paste(100," ",mu,"g/L")),
                               expression(paste(1000," ",mu,"g/L")),
                               expression(paste(10000," ",mu,"g/L")))
                    )+
  theme(legend.position=c(.8,.8))
```


# Survival
Performing the survival analysis using a Cox proportional hazards model implemented in the "survival" R package.  


## Data manipulation
Before the analysis, we removed data from after a snail had died and formatted the data for the survival analysis (a dataset of death events and censure times)

```{r Survival}
SurvData <- Data %>%
  mutate(Treatment=factor(Treatment)) %>%
  mutate(Death = case_when(Status == "alive" ~ 0, # code survival numerically
                           Status == "dead" ~ 1))%>%
  na.omit(Length) # remove entries after snail has died

# Create a data set of events and censored times (when snail survives the entire experiment)

Survival_df<- c()
for(i in 1:40){
  
  subdat <- SurvData %>% filter(Snail==i)
  
  if(length(which(subdat$Status=="dead"))!=0){ # if the snail dies, record the death event
    keep_line <- subdat[which(subdat$Status=="dead"),]}
  if(length(which(subdat$Status=="dead"))==0){ # if the snail survives, record the final sample time
    keep_line <- subdat[nrow(subdat),]}
  
  Survival_df <- rbind(Survival_df,keep_line)
}
```

## Performing the survival analysis

We performed a survival analysis using a cox proportional hazards model.  

```{r Survival Contrasts}
SurvData$TreatmentContrasts <- SurvData$Treatment # create a separate column for contrasts
overall_surv <- Surv(SurvData$Week,SurvData$Death) # create the survival object

surv_fit <- coxph(Surv(Week, Death)~Treatment, data=SurvData) 
anova(surv_fit)
pairwise_survdiff(Surv(Week, Death)~Treatment, data=SurvData)

```


## Creating the plot
```{r Survival_Fig}
Survival_plotdat <- Data %>%
  mutate(Alive = case_when(Status == "dead" ~ "0", #redoing the numeric coding so Alive is 1 
                           Status == "alive" ~ "1"),
         Alive = as.numeric(Alive)
  ) %>%
  filter(Week !=0) %>%
  group_by(Treatment,Week) %>%
  dplyr::summarize(
    Sum_alive = sum(Alive),
    Prop_alive = mean(Alive),
    SD_alive = sd(Alive),
    SE_alive = SD_alive/sqrt(length(Alive)),
    se_binom = sqrt(Prop_alive*(1-Prop_alive)/length(Alive))
  )


pd=position_dodge(.5)

surv_plot <- ggplot(Survival_plotdat, aes(x=Week,y=Prop_alive,color=Treatment))+
  geom_point(size=4,position=pd)+
  geom_errorbar(aes(ymin=Prop_alive-se_binom,ymax=Prop_alive+se_binom),position=pd,width=.2)+
  theme_classic()+
  labs(y="Proportion of snails alive")+
    scale_color_manual(values=color_pal,
                    labels = c(expression(paste(0," ",mu,"g/L")),
                               expression(paste(100," ",mu,"g/L")),
                               expression(paste(1000," ",mu,"g/L")),
                               expression(paste(10000," ",mu,"g/L")))
                    )+
  #theme(legend.position=c(.9,.8))
  theme(legend.position="none")
```


## Combined Demographics plot 
```{r combined_growthsurvivalreproduction}
library(cowplot)

p_1 <- plot_grid(growth_plot,reproduction_plot,reproduction_plot,ncol=3,labels=c("A","B","C"))

ggsave(p_1,filename="DemographicResults.eps",device="eps",width=7.5,height=3)
```


# Feeding

Feeding data was collected throughout the experiment. We were unable to quantify the amount of food, but were able to note when food was consumed or not, so the data are presented as a  binary outcome.  

## Formatting the data
```{r Feeding}
Start = as.Date("2023-02-04") #Beginning of the experiment

Feeding_data <- read.csv("Feeding_data.csv") %>% # read in feeding data and convert dates to days of the experiment
  mutate(Date = mdy(Date),
         Day = as.numeric(Date-Start),
         Fed = case_when(Food == "no" ~ 0, # changing fed to numeric
                         Food == "yes" ~ 1)
  ) %>%
  na.omit() # remove snails that died


mean(Feeding_data$Fed)
```       

## Performing the analysis
```{r}

# model & analysis
mod_feeding <- glmmTMB(Fed~Day*Treatment + (1|Snail),data=Feeding_data,family="binomial")
summary(mod_feeding)
Anova(mod_feeding)



L4 <- Feeding_data %>% filter(Treatment=="L4")
otherlevels <- Feeding_data %>% filter(Treatment!="L4")

(mean(otherlevels$Fed)-mean(L4$Fed))/(mean(otherlevels$Fed))



# model validation
simulationOutput <- simulateResiduals(fittedModel = mod_feeding, plot = T)
testDispersion(simulationOutput)

feeding_posthoc <- emmeans(mod_feeding, ~ Treatment)
pairs(feeding_posthoc)
```


## Making the plot
```{r}
Feeding_plotdat <- Feeding_data %>%
  group_by(Treatment,Day) %>%
  dplyr::summarize(
    Prop_fed = mean(Fed),
    SD_fed = sd(Fed),
    SE_fed = SD_fed/sqrt(length(Fed)),
    se_binom = sqrt(Prop_fed*(1-Prop_fed)/length(Fed))
  )

pd=position_dodge(.5)

feeding_fig <- ggplot(Feeding_plotdat, aes(x=Day,y=Prop_fed,color=Treatment))+
  geom_line()+
  geom_point(position=pd)+
  geom_errorbar(aes(ymin=Prop_fed-se_binom,ymax=Prop_fed+se_binom),position=pd,width=.2)+
  theme_classic()+
  labs(y="Proportion of snails that fed")+
  scale_color_manual(values=color_pal,
                     labels = c(expression(paste(0," ",mu,"g/L")),
                                expression(paste(100," ",mu,"g/L")),
                                expression(paste(1000," ",mu,"g/L")),
                                expression(paste(10000," ",mu,"g/L")))
  )+
  theme(legend.position = "top") +
  guides(color = guide_legend(nrow = 2))
  #theme(legend.position=c(.9,.8)) +
  
feeding_fig
ggsave(feeding_fig,filename="FeedingResults.eps",device="eps",width=3.5,height=3)


```

